Automatic theorem proving (ATP) for first-order logic (FOL) is a technology that solves a wide array of formalized tasks in domains such as mathematics and software verification.
Saturation-based automatic theorem provers, exemplified by the highly optimized Vampire prover, constitute the state of the art in ATP.

Various heuristics guide the operation of state-of-the-art provers.
Given an input problem, the chance of success may vary greatly depending on the configuration of the heuristics.
Traditionally, an expert user configures the heuristics.
In a typical prover, the number of possible configurations is large, and the response to a configuration is hard to predict.
These difficulties make machine learning (ML) an attractive alternative to relying on an expert user.

This dissertation thesis presents a line of research in ML for saturation-based ATP.
I designed and evaluated three approaches for applying ML to improve the performance of the saturation-based Vampire prover.
In the first approach, I created and evaluated a system that, given an arbitrary input problem, predicts a symbol precedence to instantiate a simplification ordering on terms.
Secondly, I created a system that predicts symbol weights for the weighted symbol-counting clause selection scheme.
Finally, my coauthors and I created a system that automatically finds complementary strategies and constructs strong strategy schedules that generalize well to unseen problems.

All of the research presented in this thesis has been published in peer-reviewed papers.
The thesis includes the papers in their final published form.
