\ctusetup{
	mainlanguage = english,
	title-english = {\glslink{ml}{Machine Learning} for Saturation-Based Theorem Proving},
	doctype = D,
	faculty = F3,
	department-english = {\href{https://cyber.felk.cvut.cz/}{Department of Cybernetics}},
	author = {\href{https://people.ciirc.cvut.cz/bartefil/}{Mgr. Filip Bártek}},
	supervisor = {\href{https://people.ciirc.cvut.cz/~urbanjo3/}{Mgr. Josef Urban, Ph.D.}},
	supervisor-address = {\href{https://www.ciirc.cvut.cz/}{Czech Institute of Informatics, Robotics and Cybernetics}\\\href{https://www.cvut.cz/}{Czech Technical University in Prague}\\Jugoslávských partyzánů 1580/3\\160 00 Praha 6\\Czech Republic},
	supervisor-specialist = {\href{https://people.ciirc.cvut.cz/~sudamar2/}{RNDr. Martin Suda, Ph.D.}},
	% https://intranet.fel.cvut.cz/cz/vv/doktorandi/obory.html
	fieldofstudy-english = {Computer Science},
	keywords-english = {\acrlong{ml}, \gls{atping}},
	keywords-czech = {\glsdisp{ml}{strojové učení}, \glsdisp{atping}{automatické dokazování}},
	% TODO: Set the correct date.
	day = 30,
	month = 9,
	year = 2024,
	specification-file = {NavrRamcovehoTematuDP_MS.pdf},
	front-list-of-figures = false,
	front-list-of-tables = false,
	pkg-hyperref = true,
}

\ctuprocess

\begin{thanks}
I thank my supervisor-specialist Martin Suda for his pragmatic guidance in research and publication.
I thank my senior supervisor Josef Urban for welcoming me in his research group and focusing my attention on \gls{ar}.
I thank the reviewers of my papers for numerous valuable suggestions that helped making the presentation of my research more accessible and complete.
I thank my colleagues, especially Jan Hůla, Thibault Gauthier, Chad Brown, and Zarathustra Goertzel, for their welcoming and supportive attitude.
I thank my partner Markéta Kociánová for her patience and care.
Finally, I thank my parents Ludmila Bártková and Jiří Bártek for the financial and emotional support they provided me in my education.
\todo[inline]{Finalize.}

\paragraph{Funding}

% https://docs.google.com/spreadsheets/d/1rHV33O5apSXxvtCM6_gqnm2dEmj-dV7ij-QQK0__85k

%\begin{itemize}
%\item This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n° 649043).
%\item This work was supported by the Grant Agency of the Czech Technical University in Prague, grant No. SGS20/215/OHK3/3T/37.
%\item The work on this paper was supported by the Czech Science Foundation grant 20-06390Y.
%\item This work was supported by the European Regional Development Fund under the Czech project AI\&Reasoning no. CZ.02.1.01/0.0/0.0/15\_003/0000466.
%\item The work on this paper was supported by the Czech Science Foundation grant 24-12759S.
%\end{itemize}

% Grants that are recent and important are prioritized.
The research presented in this doctoral thesis was supported by
the Czech Science Foundation grants 20-06390Y and 24-12759S,
the European Regional Development Fund under the Czech project AI\&Reasoning no.~CZ.02.1.01/0.0/0.0/15\_003/0000466,
% Project 649043 (AI4REASON) financed only my PAAR 2020 publication.
the \gls{erc} under the European Union's Horizon 2020 research and innovation programme (grant agreement n°~649043), and
the Grant Agency of the Czech Technical University in Prague, grant No.~SGS20/215/OHK3/3T/37.
\end{thanks}

\begin{declaration}
%Prohlašuji, že jsem předloženou práci vypracoval samostatně, a že jsem uvedl veškerou použitou literaturu.
I hereby declare I have written this doctoral thesis independently and quoted all the sources of information used in accordance with methodological instructions on ethical principles for writing an academic thesis. Moreover, I state that this thesis has neither been submitted nor accepted for any other degree.

In Prague, \today
\end{declaration}

\begin{abstract-english}
\todo[inline,color=green]{Review.}

\input{abstract.txt}

%I used the award winning prover Vampire to implement and empirically evaluate the ML designs I proposed.
%I used the FOL fragment of the TPTP problem library as the target problem domain.
%Using a problem set this broad and varied ensures a level of generality of the empirical results.
\end{abstract-english}

\begin{abstract-czech}
\todo[inline,color=red]{Translate from English.}
\end{abstract-czech}

\input{preamble-common}
\hypersetup{hidelinks}
