%\newcommand*{\IncludePaper}[2][height=\paperheight-1.5in]{\includepdf[pages=-, pagecommand={}, #1]{#2}}
\newcommand*{\IncludePaper}[2][]{}

\chapter{Results}

\todo[inline]{Consider promoting the paper sections to chapters.}

The results of my research have been published at a number of conferences and workshops.
This chapter includes a selection of these publications.
Each of the publications is self-contained.

\todo[inline]{In each section, summarize the results of the respective paper. Introduce each paper from a contemporary perspective.}

\section{Learning Precedences from Simple Symbol Features}
\label{sec:results:simple}

Filip Bártek and Martin Suda.
Learning Precedences from Simple Symbol Features.
\Acrlong{paar} 7, 2020.
\cite{DBLP:conf/cade/Bartek020}

\IncludePaper{publications/simple.pdf}

\section{Neural Precedence Recommender}
\label{sec:results:npr}

Filip Bártek and Martin Suda.
Neural Precedence Recommender.
\Acrlong{cade} 28, 2021.
\cite{DBLP:conf/cade/Bartek021}

\IncludePaper{publications/Neural Precedence Recommender.pdf}

\section{A GNN-Advised Clause Selection}
\label{sec:results:selection}

Filip Bártek and Martin Suda.
How much should this symbol weigh? A \acrshort{gnn}-Advised Clause Selection.
\Acrlong{lpar} 24, 2023.
\cite{DBLP:conf/lpar/Bartek023}

\IncludePaper{publications/weights.pdf}

\section{Regularization in Spider-Style Strategy Discovery and Schedule Construction}
\label{sec:results:regularization}

Filip Bártek, Karel Chvalovský, and Martin Suda.
Regularization in Spider-Style Strategy Discovery and Schedule Construction.
\Acrlong{ijcar} 12, 2024.
\cite{DBLP:conf/ijcar/BartekCS24}\todo{Link to the arXiv version with appendix.}

In this paper, we invented strong complementary strategies for the \gls{atper} Vampire and constructed strong schedules that generalize well.

\IncludePaper{publications/regularization.pdf}

\section{Cautious Specialization of Strategy Schedules}
\label{sec:results:cautious}

Filip Bártek, Karel Chvalovský, and Martin Suda.
Cautious Specialization of Strategy Schedules (Extended Abstract).
\Acrlong{paar} 9, 2024.
\cite{DBLP:conf/paar/BartekC024}

\IncludePaper{publications/cautious.pdf}
