Automatic theorem proving (ATP) for first-order logic (FOL) addresses a wide array of formalized tasks in domains such as mathematics and software verification. Saturation-based automatic theorem provers, notably the highly optimized Vampire prover, represent the cutting edge in ATP.

State-of-the-art provers employ various heuristics, and their success on a given problem largely depends on the configuration of these heuristics. Traditionally, configuring them has been the domain of expert users. The vast array of available configurations and the unpredictability of their outcomes complicate this task, making machine learning (ML) a promising alternative to relying only on human expertise.

This dissertation explores the application of ML in saturation-based ATP, with research focused on enhancing the performance of the Vampire prover through three developed approaches. The first approach involves a system that recommends, for an arbitrary input problem, a symbol precedence to instantiate a simplification ordering on terms in order to prune the proof search efficiently. The second approach leverages a system that recommends symbol weights for the weighted symbol-counting clause selection scheme. Finally, in collaboration with coauthors, a system was developed that autonomously finds complementary strategies (heuristic configurations) and constructs robust strategy schedules that generalize well to previously unseen problems.

All research presented in this thesis has been published in peer-reviewed articles, and the thesis features these articles in their final published forms.
